---
title: "p8105_hw6_jer2243"
author: "Juliet Rowe"
date: "2023-11-20"
output: github_document
---

## Problem 2

Load libraries
```{r}
library(tidyverse)

set.seed(1)
```


Download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Create bootstrap with 5000 samples
```{r}
bootstrap_results = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy)) |> 
  select(results) |> 
  unnest(results)

```

Extract estimates for tmin and prcp
```{r}
estimate_tmin <- bootstrap_results |> 
  filter(term == "tmin") |> 
  pull(estimate)

estimate_prcp <- bootstrap_results |>
  filter(term == "prcp") |> 
  pull(estimate)
```

Calculate r^2 and log(beta1*beta2)
```{r}
bootstrap_estimate_results <- bootstrap_results |>
  summarize(
    r_squared = (sum(estimate_tmin * estimate_prcp) / (sd(estimate_tmin) * sd(estimate_prcp)))^2,
    log_beta_product = ifelse(mean(estimate_tmin) > 0 & mean(estimate_prcp) > 0,
                              log(mean(estimate_tmin) * mean(estimate_prcp)),
                              NA)
  )

```


## Problem 3

Load and clean data

variables want to convert to factor: babysex, frace, malform, mrace
```{r}
birthweight_df <- read.csv("birthweight.csv")

birthweight_df = birthweight_df |>
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |>
  drop_na()
  

