---
title: "p8105_hw6_jer2243"
author: "Juliet Rowe"
date: "2023-11-20"
output: github_document
---

## Problem 2

Load libraries
```{r}
library(tidyverse)

set.seed(1)
```


Download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Create bootstrap with 5000 samples
```{r}
bootstrap_results = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    rsq = map(models, broom::glance)) |> 
  select(results, .id, rsq) |> 
  unnest(results) |>
  filter(term %in% c("tmin", "prcp")) |>
  group_by(.id) |>
  mutate(beta1xbeta2 = prod(estimate),
         log_beta_product = log(beta1xbeta2)) |>
  select(log_beta_product, rsq) |>
  unnest(rsq) |>
  janitor::clean_names() |>
  select(log_beta_product, id, r_squared) |>
  unique()

```

Plot distribution of log_beta_product
```{r}
bootstrap_results |>
  ggplot(aes(x=log_beta_product)) + geom_density()
```
The distribution of the bootstrap estimates of the log of the product of beta 1 and beta 2 is left skewed, unimodal, and peaks at around -5.5.


Plot distribution of r-squared
```{r}
bootstrap_results |>
  ggplot(aes(x=r_squared)) + geom_density()
```
The distribution of the bootstrap estimates of r-squared is slightly left-skewed, unimodal, and peaks at around 0.92.


Calculate 95% CI for r-squared
```{r}
bootstrap_results |>
  summarize(
    ci_lower = quantile(r_squared, 0.025),
    ci_upper = quantile(r_squared, 0.975)
  ) |>
  knitr::kable()
```


Calculate 95% CI for log beta product

## Problem 3

Load and clean data

variables want to convert to factor: babysex, frace, malform, mrace
```{r}
birthweight_df <- read.csv("birthweight.csv")

birthweight_df = birthweight_df |>
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |>
  drop_na()
  
```


