---
title: "p8105_hw6_jer2243"
author: "Juliet Rowe"
date: "2023-11-20"
output: github_document
---

## Problem 2

Load libraries
```{r}
library(tidyverse)
library(modelr)
library(mgcv)
library(dplyr)
library(knitr)

set.seed(1)
```


Download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Create bootstrap with 5000 samples
```{r}
bootstrap_results = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    rsq = map(models, broom::glance)) |> 
  select(results, .id, rsq) |> 
  unnest(results) |>
  filter(term %in% c("tmin", "prcp")) |>
  group_by(.id) |>
  mutate(beta1xbeta2 = prod(estimate),
         log_beta_product = log(beta1xbeta2)) |>
  select(log_beta_product, rsq) |>
  unnest(rsq) |>
  janitor::clean_names() |>
  select(log_beta_product, id, r_squared) |>
  unique()

```

Plot distribution of log_beta_product
```{r}
bootstrap_results |>
  ggplot(aes(x=log_beta_product)) + geom_density()
```
The distribution of the bootstrap estimates of the log of the product of beta 1 and beta 2 is left skewed, unimodal, and peaks at around -5.5.


Plot distribution of r-squared
```{r}
bootstrap_results |>
  ggplot(aes(x=r_squared)) + geom_density()
```
The distribution of the bootstrap estimates of r-squared is slightly left-skewed, unimodal, and peaks at around 0.92.


Calculate 95% CI for r-squared and log beta product
```{r}
bootstrap_CIs = bootstrap_results |>
  unique() |> ungroup() |> select(-id) |> summarize(betas_mean = mean(log_beta_product, na.rm=TRUE),
                                              betas_low = quantile(log_beta_product, 0.025, na.rm=TRUE),
                                              betas_high = quantile(log_beta_product, 0.975, na.rm=TRUE),
                                              rsq_mean = mean(r_squared),
                                              rsq_low = quantile(r_squared, 0.025),
                                              rsq_high = quantile(r_squared, 0.975))
bootstrap_CIs |> knitr::kable()
```



## Problem 3

Load and clean data

variables want to convert to factor: babysex, frace, malform, mrace
```{r}
birthweight_df <- read.csv("birthweight.csv")

birthweight_df = birthweight_df |>
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |>
  drop_na()
  
```


Proposed regression for birthweight 
```{r}
fit = lm(bwt ~ babysex + delwt + gaweeks + mrace + parity + smoken, data = birthweight_df)

summary(fit)

modelr::add_residuals(birthweight_df, fit)
modelr::add_predictions(birthweight_df, fit)
```


Plot of residuals against fitted values
```{r}
birthweight_df <- birthweight_df |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = bwt, y = resid)) + geom_violin() 

```


Comparison plots
```{r}
fit2 = lm(bwt ~ blength + gaweeks, data = birthweight_df)

fit3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*babysex*blength, data = birthweight_df)

birthweight_df |> 
  gather_predictions(fit, fit2, fit3) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = bwt, y = resid)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)

```

