---
title: "p8105_hw6_jer2243"
author: "Juliet Rowe"
date: "2023-11-20"
output: github_document
---

## Problem 2

Load libraries
```{r}
library(tidyverse)
library(modelr)
library(mgcv)
library(dplyr)
library(knitr)

set.seed(1)
```


Download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Create bootstrap with 5000 samples
```{r}
bootstrap_results = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    rsq = map(models, broom::glance)) |> 
  select(results, .id, rsq) |> 
  unnest(results) |>
  filter(term %in% c("tmin", "prcp")) |>
  group_by(.id) |>
  mutate(beta1xbeta2 = prod(estimate),
         log_beta_product = log(beta1xbeta2)) |>
  select(log_beta_product, rsq) |>
  unnest(rsq) |>
  janitor::clean_names() |>
  select(log_beta_product, id, r_squared) |>
  unique()

```

Plot distribution of log_beta_product
```{r}
bootstrap_results |>
  ggplot(aes(x=log_beta_product)) + geom_density()
```
The distribution of the bootstrap estimates of the log of the product of beta 1 and beta 2 is left skewed, unimodal, and peaks at around -5.5.


Plot distribution of r-squared
```{r}
bootstrap_results |>
  ggplot(aes(x=r_squared)) + geom_density()
```
The distribution of the bootstrap estimates of r-squared is slightly left-skewed, unimodal, and peaks at around 0.92.


Calculate number of log(beta1xbeta2) that are undefined
```{r}
nan_count <- bootstrap_results |>
  pull("log_beta_product") |>
  is.nan() |>
  sum()
```
3361 out of 5000 observations for log(beta1xbeta2) are undefined.


Calculate 95% CI for r-squared and log beta product
```{r}
bootstrap_CIs = bootstrap_results |>
  unique() |> ungroup() |> select(-id) |> summarize(betas_mean = mean(log_beta_product, na.rm=TRUE),
                                              betas_low = quantile(log_beta_product, 0.025, na.rm=TRUE),
                                              betas_high = quantile(log_beta_product, 0.975, na.rm=TRUE),
                                              rsq_mean = mean(r_squared),
                                              rsq_low = quantile(r_squared, 0.025),
                                              rsq_high = quantile(r_squared, 0.975))
bootstrap_CIs |> knitr::kable()
```
The 95% CI for log(beta1xbeta2) = (`r bootstrap_CIs$betas_low`, `r bootstrap_CIs$betas_high`).
The 95% CI for log(beta1xbeta2) = (`r bootstrap_CIs$rsq_low`, `r bootstrap_CIs$rsq_high`).


## Problem 3

Load and clean data

variables to convert to factor: babysex, frace, malform, mrace
```{r}
birthweight_df <- read.csv("birthweight.csv")

birthweight_df = birthweight_df |>
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |>
  drop_na()
  
```


Proposed regression for birthweight 
```{r}
fit = lm(bwt ~ babysex + delwt + gaweeks + mrace + parity + smoken, data = birthweight_df)

fit |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=3)

summary(fit)

```
My proposed model for birthweight includes `delwt `, mother's weight at delivery, `gaweeks`, gestational age in weeks, `mrace`, mother's race, `parity`, number of live births prior to this pregnancy, and `smoken`, average number of cigarettes smoked per day during pregnancy. I chose these variables because I know that they are all likely to have a biological effect on fetal development and therefore the birthweight of the baby when it is born. The p-values for all of these variables, aside from 0.321, are quite small so it seems like my hypothesized model was pretty good!


Plot of residuals against fitted values
```{r}
birthweight_df |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  ggplot(aes(x=resid, y=pred)) + geom_point(alpha=0.5, size=0.5)
```
The plot of residuals against fitted values appears to be clustered around 0, with the average predicted value looking to be around 3200 grams for birthweight. This appears to be a decent fit, although there are a good amount of outliers.

Cross-validation comparison
```{r}
cv_df = 
  crossv_mc(birthweight_df, 100) 

cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df |> 
  mutate(
    mod_1  = map(train, \(df) lm(bwt ~ babysex + delwt + gaweeks + mrace + parity + smoken, data = df)),
    mod_2  = map(train, \(df) lm(bwt ~ blength + gaweeks, data=df)),
    mod_3  = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data=df))) |> 
  mutate(
    rmse_1 = map2_dbl(mod_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(mod_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_3 = map2_dbl(mod_3, test, \(mod, df) rmse(model = mod, data = df)))

cv_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(), names_to="model", values_to="rmse", names_prefix="rmse_") |>
  ggplot(aes(x=model, y=rmse)) + geom_violin()
  
```

Based on the RMSE values, it looks like the 3rd model with the interaction terms is the best fit. This model has significantly lower RMSE values than the other two. 
